\documentclass[12pt]{article}
\usepackage{indentfirst}
\usepackage{fancyhdr}
\usepackage[style=apa]{biblatex}
\usepackage{filecontents}

\addbibresource{sources.bib}

\pagestyle{fancy}
\fancyhf{}
\fancyhead[R]{Bontera Research, \thepage}
\fancyhead[L]{\leftmark }
\fancyfoot[C]{Hayaan Rizvi}
\setlength{\headheight}{15pt}
\renewcommand{\footrulewidth}{0.4pt}

\begin{document}
\begin{center}
	\Huge\textbf{Bontera Research}\\

	\Large{Hayaan Rizvi}\\

	\Large{Dr. Imam}\\
	As of \today\\
	\hfill \break
	\noindent\rule{10cm}{0.4pt}

\end{center}

\setlength{\parindent}{0.5in}

\hfill \break
\newpage

\tableofcontents

\newpage
\section{Potatoes}
\subsection{Soft Rot}
\subsubsection{Description}

Bacterial soft rot affects a number of fruits and vegetables, including potatoes. It is a post-harvest disease, occuring while the crop is stored or in transit \autocite{rich2013potato}. It is characterized by a watery soft spot on the side of the crop and an strong odor.

\subsubsection{Geography and Soil Type}

Soft rot arises worldwide, from South Africa \autocite{ngadze2012pectinolytic} to the northeastern United States \autocite{ge2021genotyping}. Pérombelon notes that the soil has to be both nutritionaly deficient and over-watered for soft rot to spread. Some bacterium are specific to potato soft rot because of the cool, temperate climate where they are grown \autocite{perombelon2002potato}. Soft rot favors warm temperatures and high moisture levels in the soil and during storage \autocite{rich2013potato}.

\subsubsection{Cause}

Two main genera cause soft rot: \emph{Pectobacterium} and \emph{Dickeya} \autocite{youdkes2020potential}. Bacterial growth is accelerated by any wound or puncture of the potato skin, especially a tender spot formed by standing water or insect bites breaking the outer layer \autocite{rich2013potato}.

\subsubsection{Biosolution}

Historically, farmers have used water-management and sanitation to control soft rot, but the bacterial predator \emph{Bdellovibrio bacteriovorus} and similar organisms have shown considerable promise as well, according to a recent study by Youdkes et al. (2020). All the strains tested were effective in reducing \emph{Pectobacterium} and \emph{Dickeya} greatly. Strains introduced in the tubers before the onset of soft rot were much more effective in fighting it later on, possibly because glucose consumption by the prey was not inhibiting growth \autocite{youdkes2020potential}.

A second study found that certain varieties of rhizosphere bacteria could also be used to control soft rot caused by \emph{Pectobacterium} strains \autocite{krzyzanowska2012rhizosphere}. The results indicate that 18 various rhizobacteria were effective in inhibiting the spread of soft rot, out of the 1165 tested. The study noted, however, that the bacteria's ability to control \emph{Pectobacterium} \emph{in vitro} may not reflect the true results \emph{in vivo}.

\subsection{Golden Nematode Disease}
\subsubsection{Description}

Nematodes are nearly invisible worms that attack potato plants and tubers \autocite{rich2013potato}. Rich also reports that some nematodes cause disease directly, while others act as vectors or catalysts for viral and fungal illnesses. Plants affected by golden nematode disease have necrotic and wilting leaves with stunted growth; many do not recover, resulting in severely reduced crop yeilds \autocite{rich2013potato}.

\subsubsection{Geography and Soil Type}
The golden nematode is native to Peru, but has spread all throughout the world \autocite{rich2013potato}. They can survive in any climate that potatoes can grow, but very strict quarantine and sanitation guidelines have prevented further spread \autocite{evans1980origin}.

Mimee et al. (2015) found that optimal nematode egg hatching occurs when soil temperatures are between 59$^\circ F$ and 80$^\circ F$. The determined that increasing temperatures in cooler climates could lead to accelerated spread of golden nematode disease. Both dry weather and light soil favor the disease \autocite{rich2013potato}.

\subsubsection{Cause}

The golden nematode (\emph{Globodera rostochiensis}) is the main cause of Golden Nematode Disease \autocite{rich2013potato}. The disease gets its name from the golden or brown cysts containing nematode eggs present after an infestation \autocite{rich2013potato}. This is how the the disease spreads: the cysts cling to containers, equipment, and tubers and are transferred between fields or farms.

\subsubsection{Biosolution}

Some farmers have reported the natural decline of nematode populations because of fungi parasitism \autocite{evans1993reviews}. 10 species were isolated from the soil, \emph{Cylindrocarpon destructans} being the most promising \autocite{evans1993reviews}. That same study found that ``when an inoculum of straw colonised by \emph{C. destructans} was placed around potato seed tubers planted in [nematode] infested soil ... the numbers of juvenile stages of \emph{G. rostochiensis} ... decreased by 62\%.''


\subsection{Brown Rot}

\subsubsection{Description}
Brown rot, also known as bacterial wilt, is a very destructive bacterial disease \autocite{rich2013potato}. It was first reported it in the United States in 1896. Potatoes afflicted with brown rot excrete a ``slimy ooze'' at the base of the stem and eyes, although infected plants may also produce healthy tubers \autocite{rich2013potato}. In the early stages of the disease, the shoot system may wilt and the tubers will turn brown \autocite{kabeil2008potato}.

\subsubsection{Geography and Soil Type}
Brown rot heavily favors warmer climates; so much so that it is virtually eraticated in Canada and the northern United States \autocite{rich2013potato}. If it does infect potatoes in cooler climates, brown rot is harder to detect early on \autocite{kabeil2008potato}. The plants may not show any signs of infection until later stages of infection.

In a study by van Elsat et al. (2000), a steady decline of brown rot was observed in loamy sand soil. They also found that severe drought drastically reduced brown rot infections across the board. The disease prefers moist, slightly acid, netral, and alkaline soil \autocite{rich2013potato}.

\subsubsection{Cause}

\emph{Pseudomonas solanacearum} is the bacterium that causes brown rot \autocite{van2000survival}. There are three races, each of which attack different crops and have different optimal climates \autocite{rich2013potato}:

\textbf{Race 1}: Attacks eggplant, tobacco, tomato, and potatoes

\textbf{Race 2}: Attacks banana

\textbf{Race 3}: Attacks mainly potato, but is weakly pathogenic to tobacco

\emph{P. solanacearum} usually enters a host through a wound on the skin or roots \autocite{rich2013potato}. Rich (2013) also reports that the aflicted crops may infect the soil, transfering brown rot to any new plants grown afterwards.

\subsubsection{Biosolution}

One effective way to combat brown rot is to add sulfur to sandy soil, followed by limestone in the summer \autocite{rich2013potato}. Furthermore, a number of potato variants are resistant to \emph{P. solanacearum}; planting these cultivars will help minimize losses \autocite{rich2013potato}.

A study was also conducted by Fujiwara et al. in 2011 that experimented with using bacteriophages to control brown rot. While many of the viruses they tried were effective, application of $\mathrm{\phi}$ RSL1 was the best method. Although it did not kill all the bacterial cells, it was successful in preventing the onset of disease. $\mathrm{\phi}$ RSL1 worked much better when it was introduced before \emph{P. solanacearum} as a preventative measure, although it did also have an effect when it was introduced after \autocite{fujiwara2011biocontrol}.

\newpage


\addcontentsline{toc}{section}{References}
\printbibliography

\end{document}
